---
title: "PRDA package: Prospective and Retrospective Design Analysis."
author: 'Andreella, A., Bertoldo, G., Vesely, A., Altoe, G. and Callegher, C. '
date: "June 19, 2020"
output:
  ioslides_presentation:
    css: style.css
    incremental: yes
    logo: logo_psicostat.png
    smaller: yes
    theme: cosmo
    transition: slower
    widescreen: true
  slidy_presentation:
    incremental: yes
---

# The power ritual {data-background=background.jpg data-background-size=cover}

## The power ritual

<p class="bottom-three">
- **Replication crisis** in social sciences and medicine $\rightarrow$ Starting to <span style="color:rgb(17,57,219);">scrutinize</span> published findings by promoting large-scale replication effort.
</p>

<p class="bottom-three">
- **However**, they found smaller effects than originals $\rightarrow$ <span style="color:rgb(17,57,219);">Decline effect</span>:
  - **What does not kill statistical significance makes it stronger**;
  - **Winner curse**.
</p>

<p class="bottom-three">
- <span style="color:rgb(17,57,219);">Statistical inference</span> is too often viewed as an **isolated procedure** limited to the analysis of data that have already been collected.
</p>


<p class="bottom-three">
- From **null ritual** to **power ritual**! :(
</p>


# Design Analysis {data-background=background.jpg data-background-size=cover}

## Only power matters?

<p class="bottom-three">
 - From Gelman and Carlin (2014), effect size estimation and statistical significance are <span style="color:rgb(17,57,219);">closely related</span>:

  - **Type M (magnitude) error** (the exaggeration ratio): indicates the predictable average <span style="color:rgb(17,57,219);">overestimation</span> of an statistically significant effect;
 
  - **Type S (sign) error** (the sign error): indicates the probability to find a statistically significant effect in the <span style="color:rgb(17,57,219);">opposite direction</span> to the one considered plausible.
</p>
<p class="bottom-three">
 - Why do these errors matter? $\rightarrow$ <span style="color:rgb(17,57,219);">Hype and Disappointment Cycle!</span>
</p>
<p class="bottom-three">
 - Consider Type M and Type S errors in Design Analysis:

   - **Retrospective Design Analysis**;

   - **Prospective Design Analysis**.
</p>

# The PRDA package {data-background=background.jpg data-background-size=cover}



## Introduction
<p class="bottom-three">
How to install it?

```{r, eval = F}
devtools::load_all("C:/Users/Angela Andreella/Documents/Rpackage/PRDA_beta/R/")
```
</p>

<p class="bottom-three">
Two implementations:

 - Pearson correlation: $H_0: \rho = 0$;

 - Cohen's $d$: $H_0: d = 0$ (one sample and two samples).
</p>

## Retrospective Design Analysis

```{r, eval = F}
retrospective(sample_n1,
              effect_size,
              sample_n2 = NULL,
              effect_type = c("cohen_d","correlation"),
              alternative = c("two.sided","less","greater"))
```

<span style="color:rgb(17,57,219);">Input:</span>

**sample_n1**: sample size

**effect_size**: value of the effect size

**sample_n2**: sample size if two samples test is used 

**effect_type**: Which effect size do you want to analyze?

**alternative**: Which alternative hypothesis?

<span style="color:rgb(17,57,219);">Output:</span>

Power, Type S, Type M, and test's characteristic

## Prospective Design Analysis 

```{r, eval = F}
prospective(effect_size, 
            power, 
            effect_type = c("cohen_d","correlation"),
            alternative = c("two.sided","less","greater"))
```

<span style="color:rgb(17,57,219);">Input:</span>

**effect_size**: effect size value of the study

**power**: power value of the study 

**effect_type**: Which effect size do you want to analyze?

**alternative**: Which alternative hypothesis?

<span style="color:rgb(17,57,219);">Output:</span>

Power, Type S, Type M, and test's characteristic

# Case Studies {data-background=background.jpg data-background-size=cover}

## Pearson correlation

We consider the study published in Science by Eisenberger et al. (2003) entitled: <span style="color:rgb(17,57,219);">"Does Rejection Hurt? An fMRI Study of Social Exclusion"</span>. 
<p class="bottom-three">
 - The Anterior Cingulate Cortex (ACC) is involved in the experience of physical pain. **Could pain from social stimuli, such as social exclusion, share similar neural underpinnings?**
</p>
<p class="bottom-three"> 
 - Each of the $13$ participants play a virtual game with other two fictitious players while undergoing functional Magnetic Resonance Imaging (fMRI);
</p> 
<p class="bottom-three"> 
 - Players had to toss a virtual ball among each other in three conditions: social inclusion, explicit social exclusion and implicit social exclusion;
</p>
<p class="bottom-three">  
 - Each participant completed a self-report measure regarding their perceived distress;
</p>
<p class="bottom-three">  
 - **Correlation coefficient** between perceived distress and activity in the ACC, $r= .88$, $p < .005$.
</p>
## Pearson correlation - Retrospective Design Analysis

**Plausible Effect Size** around $\rho = .25$ (Vul and Pashler (2017))

```{r, eval = F}
retrospective(sample_n1 = 13,
              effect_size = 0.25,
              effect_type = "correlation",
              alternative = "two.sided",
              sig_level = 0.05)
```

## Pearson correlation - Prospective Design Analysis

We want the $80\%$ probability to detect an effect of at least $\rho = \pm .25$ (Vul and Pashler (2017)):

```{r, eval = F}
prospective(rho = 0.25, 
            power = 0.8, 
            effect_type = "correlation",
            alternative = "two.sided")
```

## Cohen's d

We consider the study <span style="color:rgb(17,57,219);">"A functional basis for structure-seeking: Exposure to structure promotes willingness to engage in motivated action"</span>  (Kay et al., 2014).
<p class="bottom-three">
 - $5$ studies showing that human beings have a natural tendency to perceive structure in the surrounding world.
 </p>
 <p class="bottom-three">
 - $67$ partecipants randomly assign to read two different type of text: natural phenomena are unpredictable and random or predictable and systematic.
 </p> 
 <p class="bottom-three">
 - The outcome measure was the willingness to work towards a goal that each participant chose as their most important. The expected result was higher score in structure condition than random.
 </p>
 <p class="bottom-three"> 
 - **Cohen's d** regarding the difference between the high score means under the two condition equals $0.5$, with t-test equals $2$ and p-value $0.05$.
</p>
## Cohen's d - Retrospective Design Analysis

**Plausible Effect Size** around $d = .2$

```{r, eval = F}
retrospective(sample_n1 = 34,
              sample_n2 = 33,
              effect_size = 0.2,
              effect_type = "cohen_d",
              alternative = "two.sided",
              sig_level = 0.05)
```

## Cohen's d - Prospective Design Analysis

We want the $80\%$ probability to detect an effect of at least $d = \pm .35$ (Vul and Pashler (2017)):

```{r, eval = F}
prospective(effect_size = 0.35, 
            power = 0.8, 
            effect_type = "cohen_d",
            alternative = "two.sided")
```

# Take home messages {data-background=background.jpg data-background-size=cover}

##

Design Analysis:
<p class="bottom-three">
- **Contribute to planning more robust and replicable studies**;
</p>
<p class="bottom-three">
- Use information <span style="color:rgb(17,57,219);">outside the data</span> at hand! Rather than focusing only on a single pilot or published study;
</p>
<p class="bottom-three">
- Useful as <span style="color:rgb(17,57,219);">planning</span> phase of a study and to <span style="color:rgb(17,57,219);">evaluate already conducted studies</span>;
</p>
<p class="bottom-three">
- Exploration of different <span style="color:rgb(17,57,219);">scenarios</span>;
</p>
<p class="bottom-three">
- **Further directions**: other effect sizes, Bayesian approach (Bayes Factor).
</p>
<div p style="position: fixed; bottom: 0; right: 0; width:40%; text-align: left">
Reproducible slides on https://github.com/angeella/eRum_2020.
</div>
# {#future-steps .emphasized data-background=back.jpg data-background-size=cover}

"Accept uncertainty. Be thoughtful, open, and modest.

Remember *ATOM*."

Wasserstein et al. (2019, p. 2)