---
title: "PRDA package: Prospective and Retrospective Design Analysis."
author: 'Andreella, A., Bertoldo, G., Vesely, A., Altoe, G. and Callegher, C. '
date: "June 19, 2020"
output:
  ioslides_presentation:
    css: style.css
    incremental: yes
    logo: logo_psicostat.png
    smaller: yes
    theme: cosmo
    transition: slower
    widescreen: true
  slidy_presentation:
    incremental: yes
---

# The power ritual {data-background=background.jpg data-background-size=cover}

## The power ritual

- **Replication crisis** in social sciences and medicine $\rightarrow$ Starting to <span style="color:rgb(17,57,219);">promoting</span> large-scale replication effort.
</p>

- **However**, they found smaller effects than originals $\rightarrow$ <span style="color:rgb(17,57,219);">Decline effect</span>:
  - **What does not kill statistical significance makes it stronger**;
  - **Winner curse**.

- <span style="color:rgb(17,57,219);">Statistical inference</span> is often viewed as an **isolated procedure** $\rightarrow$ from **null ritual** to **power ritual**! :(

 - From Gelman and Carlin (2014), effect size estimation and statistical significance are <span style="color:rgb(17,57,219);">closely related</span>:

   1. **Type M (magnitude) error** (the exaggeration ratio): indicates the predictable average <span style="color:rgb(17,57,219);">overestimation</span> of an statistically significant effect;
 
   2. **Type S (sign) error** (the sign error): indicates the probability to find a statistically significant effect in the <span style="color:rgb(17,57,219);">opposite direction</span> to the one considered plausible.

 - Design Analysis: **Retrospective** and  **Prospective**






# The PRDA package {data-background=background.jpg data-background-size=cover}



## Introduction
<p class="bottom-three">
How to install it?

```{r, warning=FALSE,eval = F}
devtools::install_github("ClaudioZandonella/PRDA_beta")
```
</p>

<p class="bottom-three">
Two implementations:

 - **Pearson correlation**: $H_0: \rho = 0$;

 - **Cohen's d**: $H_0: d = 0$ (one sample and two samples).
</p>

## Retrospective Design Analysis

```{r, eval = F}
retrospective(sample_n1, effect_size, sample_n2 = NULL,
              effect_type = c("cohen_d","correlation"),
              alternative = c("two.sided","less","greater"),
              tl = -Inf, tu = Inf)
```

<span style="color:rgb(17,57,219);">**INPUT**:</span>

 - ``sample_n1``: sample size

 - ``effect_size``: value of the effect size or function indicating the hypothetical population effect size.

 - ``sample_n2``: sample size if two samples test is used 

 - ``effect_type``: Which effect size do you want to analyze?

 - ``alternative``: Which alternative hypothesis?

<span style="color:rgb(17,57,219);">**OUTPUT**:</span>  

 - ``power``, ``typeM``, and ``typeS`` and some ``test_info``

## Prospective Design Analysis 

```{r, eval = F}
prospective(effect_size, power, ratio_n = 1,
            effect_type = c("cohen_d","correlation"),
            alternative = c("two.sided","less","greater"),
            tl = -Inf, tu = Inf)
```

<span style="color:rgb(17,57,219);">**INPUT**:</span>

 - ``effect_size``: effect size value of the study or function indicating the hypothetical population effect size

 - ``power``: power value of the study 

 - ``ratio_n``: ratio between ``sample_n1`` and ``sample_n2``

 - ``effect_type``: Which effect size do you want to analyze?

 - ``alternative``: Which alternative hypothesis?

<span style="color:rgb(17,57,219);">**OUTPUT**:</span>

 - ``power``, ``typeM``, and ``typeS`` and some ``test_info``

# Case Studies {data-background=background.jpg data-background-size=cover}

## Pearson correlation

We consider the study by Eisenberger et al. (2003) entitled: <span style="color:rgb(17,57,219);">"Does Rejection Hurt? An fMRI Study of Social Exclusion"</span>. 
<p class="bottom-three">
 - The Anterior Cingulate Cortex (ACC) is involved in the experience of physical pain. **Could pain from social stimuli, such as social exclusion, share similar neural underpinnings?**
</p>
<p class="bottom-three"> 
 - Each of the $13$ participants play a virtual game with other **two fictitious players** while undergoing functional Magnetic Resonance Imaging (fMRI);
</p> 
<p class="bottom-three"> 
 - Players had to toss a virtual ball among each other in three conditions: **social inclusion**, **explicit social exclusion** and **implicit social exclusion**;
</p>
<p class="bottom-three">  
 - Each participant completed a **self-report measure** regarding their perceived distress;
</p>
<p class="bottom-three">  
 - **Correlation coefficient** between perceived distress and activity in the ACC, $r= .88$, $p < .005$.
</p>
## Pearson correlation - Retrospective Design Analysis

We consider a **plausible effect size** equals $\rho = .25$ (Vul and Pashler, 2017):

```{r, eval = F}
retrospective(sample_n1 = 13,
              effect_size = 0.25,
              effect_type = "correlation",
              alternative = "two.sided",
              sig_level = 0.05)
```

## Pearson correlation - Prospective Design Analysis

We want the $80\%$ probability to detect a **plausible effect size** of at least $\rho = .25$ (Vul and Pashler, 2017):

```{r, eval = F}
prospective(effect_size = 0.25, 
            power = 0.8, 
            effect_type = "correlation",
            alternative = "two.sided")
```

## Cohen's d

We consider the study by Kay et al., 2014 entitled <span style="color:rgb(17,57,219);">"A functional basis for structure-seeking: Exposure to structure promotes willingness to engage in motivated action"</span>.
 <p class="bottom-three">
 - $67$ partecipants randomly assign to read two different type of text: natural phenomena are  **unpredictable and random ** or  **predictable and systematic**.
 </p> 
 <p class="bottom-three">
 - The outcome measure was the inclination to work towards a goal that each participant chose as their most important. The expected result was **higher score in structure condition than random**.
 </p>
 <p class="bottom-three"> 
 - **Cohen's d** regarding the difference between the high score means under the two condition equals $0.5$, with t-test equals $2$ and p-value $0.05$.
</p>
## Cohen's d - Retrospective Design Analysis

We consider a **plausible effect size** equals $d = .35$ (Open Science Collaboration, 2015):

```{r, eval = F}
retrospective(sample_n1 = 34,
              sample_n2 = 33,
              effect_size = 0.2,
              effect_type = "cohen_d",
              alternative = "two.sided",
              sig_level = 0.05)
```

## Cohen's d - Prospective Design Analysis

We want the $80\%$ probability to detect a **plausible effect size** of at least $d = .35$ (Open Science Collaboration, 2015):

```{r, eval = F}
prospective(effect_size = 0.35, 
            power = 0.8, 
            effect_type = "cohen_d",
            alternative = "two.sided")
```

## Cohen's d - Prospective Design Analysis

We can also explore inferential risk:

```{r, eval = F}
prospective(effect_size = 0.35, 
            power = 0.6, 
            effect_type = "cohen_d",
            alternative = "two.sided")
```



# Take home messages {data-background=background.jpg data-background-size=cover}

##

Design Analysis:
<p class="bottom-three">
- **Type M** and **Type S** errors quantify the inferential risks in terms of effect size estimation $\rightarrow$ <span style="color:rgb(17,57,219);">surpass power ritual</span>
</p>
<p class="bottom-three">
- Contribute to planning more robust and replicable studies;
</p>
<p class="bottom-three">
- Use information <span style="color:rgb(17,57,219);">outside the data</span> at hand! Rather than focusing only on a single pilot or published study;
</p>
<p class="bottom-three">
- Contribute <span style="color:rgb(17,57,219);">planning</span> to planning more robust and replicable studies (**PROSPECTIVE**) and to <span style="color:rgb(17,57,219);">evaluate already conducted studies (**RETROSPECTIVE**)</span>;
</p>
<p class="bottom-three">
- Exploration of different <span style="color:rgb(17,57,219);">scenarios</span>;
</p>
<p class="bottom-three">
- **Further directions**: other effect sizes, Bayesian approach (Bayes Factor).
</p>
<div p style="position: fixed; bottom: 0; right: 0; width:40%; text-align: left">
Slides and articles on https://github.com/angeella/eRum_2020.
</div>
# {#future-steps .emphasized data-background=back.jpg data-background-size=cover}

"Accept uncertainty. Be thoughtful, open, and modest.

Remember *ATOM*."

Wasserstein et al. (2019, p. 2)
